---
title: "Geog 509 Chapter 2 Problem Set"
author: "Jesse Piburn"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2E1
2 and 4 are both correct
$\Pr(rain|monday)$ and $\frac{\Pr(rain|monday)}{\Pr(monday)}$

## 2M1
```{r 2m1-grid-approx, warning=FALSE, message=FALSE}
library(tidyverse)

tibble(
  label = c("W, W, W", "W, W, W, L", "L, W, W, L, W, W, W"),
      w = c(3, 3, 5),
      n = c(3, 4, 7),
      prior  = 1
    ) %>% 
  expand(
    nesting(w, n, prior, label),
    p_grid = seq(from = 0, to = 1, length.out = 30)
  ) %>%
  group_by(w, n) %>%
  mutate(
    likelihood = dbinom(w, size = n, prob = p_grid),
    unstd_posterior = likelihood * prior,             
    posterior = unstd_posterior / sum(unstd_posterior)
  ) %>% 
  ggplot(aes(x = p_grid, y = posterior)) +
  geom_point() +
  geom_line() +
  labs(
    subtitle = "Grid Approximated Posterior Distrubtions",
    x = "probability of water",
    y = "posterior probability") +
  theme(panel.grid = element_blank()) +
facet_wrap(~label)
```



## 2H1

\[\Pr(twins|A)=0.1\]

\[\Pr(twins|B)=0.2\]

\[\Pr(A)=0.5\]

\[\Pr(B)=0.5\]

\[\Pr(twins)=\Pr(twins|A)\Pr(A)+\Pr(twins|B)\Pr(B)=0.1(0.5)+0.2(0.5)=0.15\]

\[
\Pr(A|twins)=\frac{\Pr(twins|A)\Pr(A)}{\Pr(twins)}=\frac{0.1(0.5)}{0.15}=\frac{1}{3}
\]

\[
\Pr(B|twins)=\frac{\Pr(twins|B)\Pr(B)}{\Pr(twins)}=\frac{0.2(0.5)}{0.15}=\frac{2}{3}
\]

\[
\Pr(twins)=\Pr(twins|A)\Pr(A)+\Pr(twins|B)\Pr(B)=0.1(\frac{1}{3})+0.2(\frac{2}{3})=\frac{1}{6}
\]

## 2H2
\[
\Pr(A|twins)=\frac{\Pr(twins|A)\Pr(A)}{\Pr(twins)}=\frac{0.1(0.5)}{0.15}=\frac{1}{3}
\]


## 2H3
\[
\Pr(single|A)=1–\Pr(twins|A)=1–0.1=0.9
\]

\[
\Pr(single|B)=1–\Pr(twins|B)=1–0.2=0.8
\]

\[
\Pr(A)=\frac{1}{3}
\]

\[
\Pr(B)=\frac{2}{3}
\]

\[
\Pr(single)=\Pr(single|A)\Pr(A)+\Pr(single|B)\Pr(B)=0.9(\frac{1}{3})+0.8(\frac{2}{3})=\frac{5}{6}
\]

\[
\Pr(A|single)=\frac{\Pr(single|A)\Pr(A)}{\Pr(single)}=\frac{0.9(1/3)}{5/6}=0.36
\]
