---
title: |
    | Geog 509: Bayesian Data Analysis
    | Chapter 4 Problem Set
author: "Jesse Piburn"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 4E1
\[
y_i \sim \mathrm{Normal}(\mu, \sigma) 
\]

## 4E2
There are two parameters $\mu$ and $\sigma$

## 4E3

\[
\Pr(\mu,\sigma|y) = \frac{\prod_i \mathrm{Normal} (y_i|\mu,\sigma) \mathrm{Normal} (\mu|0,10) \mathrm{Uniform}(\sigma|0,10)}{\int \int \prod_i \mathrm{Normal}(h_i|\mu,\sigma) \mathrm{Normal}(\mu|0,10) \mathrm{Uniform}(\sigma|0,10)d\mu d\sigma}
\]

## 4M1
```{r 4m1, warning=FALSE, message=FALSE}
n <- 10000
mu <- rnorm(n, 0, 10)
sigma <- runif(n, 0, 10)
y_prior <- rnorm(n, mu, sigma)
hist(y_prior)
```

## 4M4
\[
h_i \sim \mathrm{Normal}(\mu,\sigma) \\
\mu_i = \alpha + \beta t_i \\
\alpha \sim \mathrm{Normal}(160, 50) \\
\beta \sim \mathrm{Normal}(0, 10) \\
\sigma \sim \mathrm{Uniform}(0, 50)
\]

## 4M5
\[
h_i \sim \mathrm{Normal}(\mu,\sigma) \\
\mu_i = \alpha + \beta t_i \\
\alpha \sim \mathrm{Normal}(120, 50) \\
\beta \sim \mathrm{Uniform}(0, 30) \\
\sigma \sim \mathrm{Uniform}(0, 50)
\]

Changed the mean for average height and also changed the distribution of of $\beta$ to be uniform to restrict it be greater than zero

## 4M6
The variance is the square of $\sigma$ so we can adjust the the following

\[
\sigma \sim \mathrm{Uniform}(0, 8)
\]

## 4H1
```{r 4h1, message=FALSE, warning=FALSE}
library(rethinking)
library(tidyverse)
library(brms)

data(Howell1)
d <- Howell1
rm(Howell1)
detach(package:rethinking, unload = T)

d2 <- 
  d %>%
  filter(age >= 18)

b4.3 <- 
  brm(data = d2, family = gaussian,
      height ~ 1 + weight,
      prior = c(prior(normal(156, 100), class = Intercept),
                prior(normal(0, 10), class = b),
                prior(uniform(0, 50), class = sigma)),
      iter = 41000, warmup = 40000, chains = 4, cores = 4,
      seed = 4)


new_weight <- tibble(weight = c(46.95, 43.72, 64.78, 32.59, 54.63))

predict_df <- 
  predict(
    b4.3,
    newdata = new_weight
   ) %>%
  as_tibble() %>%
  bind_cols(new_weight)
```
```{r, echo=FALSE}
knitr::kable(predict_df)
```




