---
title: |
    | Geog 509: Bayesian Data Analysis
    | Chapter 5 Problem Set
author: "Jesse Piburn"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




## 5E1  - 5E4

### 5E1
Models 2,3, and 4 are multiple regressions

### 5E2
$l_i = latitude$, $A_i = Animal Diversity$, and $P_i = Plant Diversity$ 

\[
\begin{aligned}
l_i &\sim \mathrm{Normal}(\mu_i,\sigma) \\
\mu_i &= \alpha + \beta_A A_i + \beta_P P_i \\
\alpha &\sim \mathrm{Normal}(0, 50) \\
\beta_A &\sim \mathrm{Normal}(0, 20) \\
\beta_P &\sim \mathrm{Normal}(0, 20) \\
\sigma &\sim \mathrm{Uniform}(0, 50)
\end{aligned}
\]

### 5E3
$t_i =$ Time to PhD, $F_i =$ amount of funding, and $S_i =$ laboratory size 

\[
\begin{aligned}
t_i &\sim \mathrm{Normal}(\mu_i,\sigma) \\
\mu_i &= \alpha + \beta_F F_i + \beta_S S_i \\
\alpha &\sim \mathrm{Normal}(0, 50) \\
\beta_F &\sim \mathrm{Normal}(0, 20) \\
\beta_S &\sim \mathrm{Normal}(0, 20) \\
\sigma &\sim \mathrm{Uniform}(0, 50)
\end{aligned}
\]

Assuming "positively associated" in the context of the question means recieving the degree faster
is a positive outcome, that would mean as funding and lab size increase, time to degree would decrease.
This would result in both funding and lab size having negative slopes.


### 5E4
Models 1, 3, 4 and 5 are inferentially equivalent


### 5M3
High divorce rate can cause a higher marriage rate in the sense that a marriage rate 
counts the number of marriages per some fixed amount of people. If people are getting 
divorced at a higher rate, some of those people are likely to get married again, therefore
inreasing the number of marriages while the count of people remains the same. One potential
way to evaluate this relationship is the regress marriage rate on both divorce rate and some
indicator of remarriage rate.



### 5M5
In this model we could include indicators for each of the causes in the hypotheses.
For example, higher gas prices could lead to less driving and therefore more exercise, means
that we would want to include measures for gas price, driving (e.g. miles driven per week),
and exercise (e.g. time exercising per week). Including the claims in the second hypothesis
we would want to add a measure of how often people eat out at restaurants (e.g. times per week) and also a measure
of how much they eat at the restaurant (e.g. calories per meal... assuming you could actually measure that). 



## 5H1-5H3
Things we'll need for all the questions
```{r 5h-prep, message=FALSE, warning=FALSE}
library(brms)
library(tidyverse)
library(tidybayes)
library(bayesplot)
library(patchwork)


data(foxes, package = "rethinking")
d <- foxes

rm(foxes)
```


*note*: All of code for for generating the charts was omitted from this report as it was
distracting and took up too much space. If interested, please see the source Rmd [here](https://github.com/jpiburn/geog509/blob/master/homework/ch5.Rmd)

 
### 5H1

#### weight ~ area
```{r h1.1-model, message=FALSE, cache=TRUE}
# weight ~ area
h1.area <- 
  brm(data = d, family = gaussian,
      weight ~ 1 + area,
      prior = c(
        prior(normal(0, 10), class = Intercept),
        prior(normal(0, 5), class = b),
        prior(cauchy(0, 1), class = sigma)
        ),
      iter = 2000, warmup = 500, chains = 4, cores = 4,
      seed = 1988)

h1.area
```

```{r, echo=FALSE, message=FALSE, fig.height=1.5, fig.width=5}
h1.area.plot <- 
posterior_samples(h1.area) %>% 
  select(
    -lp__, 
    -Intercept
    ) %>% 
  gather() %>% 
  ggplot(
    aes(x = value, y = reorder(key, value))
    ) +
  geom_vline(
    xintercept = 0, 
    color = "firebrick4", 
    alpha = 1/10
    ) +
  stat_pointintervalh(
    point_interval = mode_hdi, 
    .width = .95, 
    size = 3/4, 
    color = "firebrick4"
    ) +
  labs(
    title = "weight ~ area",
    x = NULL, 
    y = NULL
    ) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    panel.grid.major.y = element_line(
      color = alpha("firebrick4", 1/4), 
      linetype = 3
      ),
    axis.text.y  = element_text(hjust = 0),
    axis.ticks.y = element_blank())
```


#### weight ~  groupsize
```{r, message=FALSE, cache=TRUE}
# weight ~ groupsize
h1.groupsize <- 
  brm(data = d, family = gaussian,
      weight ~ 1 + groupsize,
      prior = c(
        prior(normal(0, 10), class = Intercept),
        prior(normal(0, 5), class = b),
        prior(cauchy(0, 1), class = sigma)
        ),
      iter = 2000, warmup = 500, chains = 4, cores = 4,
      seed = 1988)

h1.groupsize
```

```{r, echo=FALSE, message=FALSE, fig.height=1.5, fig.width=5}
h1.groupsize.plot <- 
posterior_samples(h1.groupsize) %>% 
  select(
    -lp__, 
    -Intercept
    ) %>% 
  gather() %>% 
  ggplot(
    aes(x = value, y = reorder(key, value))
    ) +
  geom_vline(
    xintercept = 0, 
    color = "firebrick4", 
    alpha = 1/10
    ) +
  stat_pointintervalh(
    point_interval = mode_hdi, 
    .width = .95, 
    size = 3/4, 
    color = "firebrick4"
    ) +
  labs(
    title = "weight ~ groupsize",
    x = NULL, 
    y = NULL
    ) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    panel.grid.major.y = element_line(
      color = alpha("firebrick4", 1/4), 
      linetype = 3
      ),
    axis.text.y  = element_text(hjust = 0),
    axis.ticks.y = element_blank())
```


```{r, echo=FALSE}
f1.area <- 
  fitted(h1.area) %>%
  as_tibble() %>%
  bind_cols(d)

p1.area <- 
ggplot(
  data = f1.area,
  aes(x = area, y = Estimate)
  ) +
  geom_smooth(
    aes(ymin = Q2.5, ymax = Q97.5),
    stat = "identity",
    fill = "firebrick", 
    color = "firebrick4", 
    alpha = 1/5, 
    size = 1/4
    ) +
  geom_point(
    aes(y = weight), 
    size = 2, 
    color = "firebrick4"
    ) +
  labs(
    title = NULL,
    x = "Area",
    y = "Weight"
    ) +
  coord_cartesian(
    xlim = range(f1.area$area), 
    ylim = range(f1.area$weight)
    ) +
  theme_bw() +
  theme(
    panel.grid = element_blank()
    )  
```

```{r, echo=FALSE}
f1.groupsize <- 
  fitted(h1.groupsize) %>%
  as_tibble() %>%
  bind_cols(d)

p1.groupsize <- 
ggplot(
  data = f1.groupsize,
  aes(x = groupsize, y = Estimate)
  ) +
  geom_smooth(
    aes(ymin = Q2.5, ymax = Q97.5),
    stat = "identity",
    fill = "firebrick", 
    color = "firebrick4", 
    alpha = 1/5, 
    size = 1/4
    ) +
  geom_point(
    aes(y = weight), 
    size = 2, 
    color = "firebrick4"
    ) +
  labs(
    title = NULL,
    x = "Group Size",
    y = NULL
    ) +
  coord_cartesian(
    xlim = range(f1.groupsize$groupsize), 
    ylim = range(f1.groupsize$weight)
    ) +
  theme_bw() +
  theme(
    panel.grid = element_blank()
    )  
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5}

area_plots <- 
(h1.area.plot) /
(p1.area) + 
  plot_layout(
    heights = c(1, 3)
    )

groupsize_plots <- 
(h1.groupsize.plot) /
(p1.groupsize) + 
  plot_layout(
    heights = c(1, 3)
    )


(area_plots | groupsize_plots) +
  plot_annotation(
    title = paste(
      "When considered individually, groupsize has a small negative",
      "relationship with weight\nwhile area has little to none"
      )
    )
```


## 5H2
#### weight ~ area + groupsize
```{r h2.1-model, message=FALSE, cache=TRUE}
# weight ~ area + groupsize
h2 <- 
  brm(data = d, family = gaussian,
      weight ~ 1 + area + groupsize,
      prior = c(
        prior(normal(0, 10), class = Intercept),
        prior(normal(0, 5), class = b),
        prior(cauchy(0, 1), class = sigma)
        ),
      iter = 2000, warmup = 500, chains = 4, cores = 4,
      seed = 1988)
h2
```


```{r, echo=FALSE, message=FALSE, fig.height=1.5, fig.width=5}
h2.plot <- 
posterior_samples(h2) %>% 
  select(
    -lp__, 
    -Intercept
    ) %>% 
  gather() %>% 
  ggplot(
    aes(x = value, y = reorder(key, value))
    ) +
  geom_vline(
    xintercept = 0, 
    color = "firebrick4", 
    alpha = 1/10
    ) +
  stat_pointintervalh(
    point_interval = mode_hdi, 
    .width = .95, 
    size = 3/4, 
    color = "firebrick4"
    ) +
  labs(
    title = "weight ~ area + groupsize",
    x = NULL, 
    y = NULL
    ) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    panel.grid.major.y = element_line(
      color = alpha("firebrick4", 1/4), 
      linetype = 3
      ),
    axis.text.y  = element_text(hjust = 0),
    axis.ticks.y = element_blank())
```

```{r, echo=FALSE}
nd <- 
  tibble(
   area = seq(
      from = floor(min(d$area)), 
      to   = ceiling(max(d$area)), 
      length.out = 30
      ),
   groupsize = mean(d$groupsize)
    )

h2_mean_groupsize_plot <- 
fitted(h2, newdata = nd) %>% 
  as_tibble() %>% 
  rename(
    f_ll = Q2.5,
    f_ul = Q97.5
    ) %>% 
  bind_cols(
    predict(h2, newdata = nd) %>% 
      as_tibble() %>% 
      transmute(
        p_ll = Q2.5,
        p_ul = Q97.5
        ),
    nd) %>% 
  ggplot(
    aes(x = area, y = Estimate)
    ) +
  geom_ribbon(
    aes(ymin = p_ll, ymax = p_ul),
    fill = "firebrick", 
    alpha = 1/5
    ) +
  geom_smooth(
    aes(ymin = f_ll, ymax = f_ul),
    stat = "identity",
    fill = "firebrick", 
    color = "firebrick4", 
    alpha = 1/5, 
    size = 1/4
    ) +
  coord_cartesian(
    xlim = range(d$area), 
    ylim = c(0, 8)
    ) +
  labs(
    title = "Area vs Weight",
    subtitle = paste0(
      "Holding Groupsize contstant at its mean of ",
      round(mean(d$groupsize), 2)
      ),
    y = NULL) + # NULL becasue we are going to combine the plots
  theme_bw() +
  theme(
    panel.grid = element_blank()
    )  
```

```{r, echo=FALSE}
nd <- 
  tibble(
   groupsize = seq(
      from = floor(min(d$groupsize)), 
      to   = ceiling(max(d$groupsize)), 
      length.out = 30
      ),
   area = mean(d$area)
    )

h2_mean_area_plot <- 
fitted(h2, newdata = nd) %>% 
  as_tibble() %>% 
  rename(
    f_ll = Q2.5,
    f_ul = Q97.5
    ) %>% 
  bind_cols(
    predict(h2, newdata = nd) %>% 
      as_tibble() %>% 
      transmute(
        p_ll = Q2.5,
        p_ul = Q97.5
        ),
    nd) %>% 
  ggplot(
    aes(x = groupsize, y = Estimate)
    ) +
  geom_ribbon(
    aes(ymin = p_ll, ymax = p_ul),
    fill = "firebrick", 
    alpha = 1/5
    ) +
  geom_smooth(
    aes(ymin = f_ll, ymax = f_ul),
    stat = "identity",
    fill = "firebrick", 
    color = "firebrick4", 
    alpha = 1/5, 
    size = 1/4
    ) +
  coord_cartesian(
    xlim = range(d$groupsize), 
    ylim = c(0, 8)
    ) +
  labs(
    title = "Groupsize vs Weight",
    subtitle = paste0(
      "Holding Area contstant at its mean of ",
      round(mean(d$area), 2)
      ),
    y = "Weight") +
  theme_bw() +
  theme(
    panel.grid = element_blank()
    )  
```

```{r, echo=FALSE, fig.width=8, fig.height=6}

h2.plot / (h2_mean_area_plot + h2_mean_groupsize_plot) +
  plot_layout(
    heights = c(1, 3)
    ) +
  plot_annotation(
    title = "When considered together however, the effect of both are intensified",
    subtitle = "Area now has a very strong positive effect when accounting for groupsize"
    )

```

The results here are different than when considering them invididually because we are
asking the model a different question and when we ask a different question we shouldn't be
surprised when we get different answers. When considering only `weight ~ area`, for example.
We are asking, what is the predictive power of area for predicting weight? However, when
we include groupsize as well, the question becomes what is the predictive power of area, once I
already know groupsize.


## 5H3
#### weight ~ avgfood + groupsize
```{r h3.ag-model, message=FALSE, cache=TRUE}
# weight ~ avgfood + groupsize
h3.ag <- 
  brm(data = d, family = gaussian,
      weight ~ 1 + avgfood + groupsize,
      prior = c(
        prior(normal(0, 10), class = Intercept),
        prior(normal(0, 5), class = b),
        prior(cauchy(0, 1), class = sigma)
        ),
      iter = 2000, warmup = 500, chains = 4, cores = 4,
      seed = 1988)

h3.ag
```

```{r, echo=FALSE, message=FALSE, fig.height=1.5, fig.width=5}
h3.ag.plot <- 
posterior_samples(h3.ag) %>% 
  select(
    -lp__, 
    -Intercept
    ) %>% 
  gather() %>% 
  ggplot(
    aes(x = value, y = reorder(key, value))
    ) +
  geom_vline(
    xintercept = 0, 
    color = "firebrick4", 
    alpha = 1/10
    ) +
  stat_pointintervalh(
    point_interval = mode_hdi, 
    .width = .95, 
    size = 3/4, 
    color = "firebrick4"
    ) +
  labs(
    title = "weight ~ avgfood + groupsize",
    x = NULL, 
    y = NULL
    ) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    panel.grid.major.y = element_line(
      color = alpha("firebrick4", 1/4), 
      linetype = 3
      ),
    axis.text.y  = element_text(hjust = 0),
    axis.ticks.y = element_blank())
```

#### weight ~ avgfood + groupsize + area
```{r h3.aga-model, message=FALSE, cache=TRUE}
# weight ~ avgfood + groupsize + area
h3.aga <- 
  brm(data = d, family = gaussian,
      weight ~ 1 + avgfood + groupsize + area,
      prior = c(
        prior(normal(0, 10), class = Intercept),
        prior(normal(0, 5), class = b),
        prior(cauchy(0, 1), class = sigma)
        ),
      iter = 2000, warmup = 500, chains = 4, cores = 4,
      seed = 1988)

h3.aga
```

```{r, echo=FALSE, message=FALSE, fig.height=1.5, fig.width=5}
h3.aga.plot <- 
posterior_samples(h3.aga) %>% 
  select(
    -lp__, 
    -Intercept
    ) %>% 
  gather() %>% 
  ggplot(
    aes(x = value, y = reorder(key, value))
    ) +
  geom_vline(
    xintercept = 0, 
    color = "firebrick4", 
    alpha = 1/10
    ) +
  stat_pointintervalh(
    point_interval = mode_hdi, 
    .width = .95, 
    size = 3/4, 
    color = "firebrick4"
    ) +
  labs(
    title = "weight ~ avgfood + groupsize + area",
    x = NULL, 
    y = NULL
    ) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    panel.grid.major.y = element_line(
      color = alpha("firebrick4", 1/4), 
      linetype = 3
      ),
    axis.text.y  = element_text(hjust = 0),
    axis.ticks.y = element_blank())
```


```{r, echo=FALSE, fig.height=8, fig.width=5, warning=FALSE, message=FALSE}
layout <- "
A
B
C
D
E
"

h1.area.plot + 
  h1.groupsize.plot + 
  h2.plot + 
  h3.ag.plot +
  h3.aga.plot + 
  plot_layout(design = layout) +
  plot_annotation(
    title = "Comparison of weight prediction models"
    ) &
  xlim(c(-1,6.1))
```

**(a)**: `area` is a better predictor of `weight` than `avgfood` for two reasons.
1) The amount of food available is a consequence of the area and 
2) avgfood has very large credible intervals even when area is not included in the model because
it has a very high correlation with both groupsize and area as seen below
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=6}
library(GGally)

ggpairs(d)

```

&nbsp;

**(b)**: This occurs because of weak identifiability in the model caused by multicollinearity between area and avgfood 